### 2019-05-31 Paper/main.tex
no message

\documentclass{modified}
\copyrightyear{2017} \pubyear{2017}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Manuscript Category}

\setcitestyle{square,sort,comma,numbers}

\usepackage[colorlinks,
            linkcolor=blue,
            anchorcolor=blue,
            citecolor=blue,
			urlcolor=blue
            ]{hyperref}

\usepackage{url}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{color}

\newcommand{\comR}[1]{#1}
\newcommand{\comG}[1]{#1}
\newcommand{\comB}[1]{#1}
\newcommand{\comM}[1]{#1}
\newcommand{\comC}[1]{#1}

\begin{document}
\firstpage{1}

\subtitle{Subject Section}

\title[Bearing Fault Diagnosis Based on Deep Belief Network]{Bearing Fault Diagnosis Based on Deep Belief Network}
\author[Sun.]{Wujie Sun\,$^{\text{\sfb 1,}*}$}
\address{$^{\text{\sf 1}}$School of Software Engineering, South China University of Technology, Guangzhou 510006, P.R. China.}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{%\textbf{Motivation:} 
It is important to judge whether the bearing is faulty and the possible fault location based on the acceleration of the bearing. However, the acceleration of the bearing is affected by noise, and it is difficult to rely on observation to find the law. This problem can be solved by using wavelet packet decomposition and deep belief network. In this paper, we use wavelet packet decomposition for noise reduction, and extract features from the noise-reduced data, and then use the deep belief network for training. The experimental results show its effectiveness.\\
\textbf{Keywords:} Deep belief network, fault diagnosis, wavelet packet decomposition, feature extraction\\
\textbf{Contact:} \href{wjsunscut@163.edu.cn}{wjsunscut@163.edu.cn}
}

\maketitle

\section{Introduction}

\section{Delief Belief Network}

\section{Experiments}

\section{Conclusions}

%\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
\bibliographystyle{unsrt}
%\bibliographystyle{bioinformatics}

%\bibliographystyle{plain}
%
%\bibliography{Document}

%\bibliographystyle{natbib}
%\bibliographystyle{myunsrt} 
\bibliography{ref}

\end{document}